---
title: "Project 3: College-level Analysis"
author: "Madison Corinha"
date: "May 10, 2020"
output: html_document
---

```{r, warning = FALSE, message = FALSE}
# Packages! (oh boy there's going to be a lot of these)
library(tidyverse)
library(tidymodels)
library(ggplot2)
library(ggridges)
```


```{r, echo = FALSE, warning = FALSE, message = FALSE}
# Data! (TidyTuesday 2020-03-10)
tuition_cost <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/tuition_cost.csv')

tuition_income <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/tuition_income.csv') 

salary_potential <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/salary_potential.csv')

historical_tuition <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/historical_tuition.csv')

diversity_school <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/diversity_school.csv')
```


MATH 228: Introduction to Data Science is a college course, and the work for this course (at the very least) must match the quality of work generally expected from an undergraduate-level student. For this project, you could say that a "college-level analysis" is required. What better data is there to perform a *"college-level analysis"* with than data *actually about colleges*!

You can find the TidyTuesday page for these data [here](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-03-10/readme.md).

General research questions:

  - Does one state have a higher avaerage tuition than any other states?
  - Is there an association between type of school (public, private, for-profit) and price of room and board?
  - Can we predict price of room and board from the available data?
  - Are there any trends in the difference between total cost for tuition and net cost for tuition (average actually paid after scholarships/awards)?
  - Is there an association between percent of students enrolled that are women and tuition?
 
  - Is tuition for colleges in Boston significantly different than tuition for colleges in Massachusetts?
  - seven sisters?



Let's start with this question: Does one state have a higher in-state tuition than any other states?

```{r}
tuition_cost %>% 
  filter(degree_length == "4 Year") %>% # only looking at 4 year colleges
  group_by(state) %>% 
  summarise(av_in_state_tu = mean(in_state_tuition)) %>% # creating variable for average in-state tuition for whole state
  na.omit() %>% # get rid of empty/NA state values
  ggplot(aes(x = fct_reorder(state, av_in_state_tu), y = av_in_state_tu)) + # fct_reorder() organizes bars in descending order (it looks WAY better!)
  geom_col() +
  theme_bw() + # my favorite theme right after theme_void()
  coord_flip() +
  labs(x = "", y = "", # x and y axis labels are redundant here
       title = "Average In-State Tuition by State",
       subtitle = "For 4-year degree programs",
       caption = "Source: Tuitiontracker.org") +
  scale_y_continuous(labels = scales::dollar_format()) # makes y axis labels into $ with commas
```









