---
title: "Project 3: College-level Analysis"
author: "Madison Corinha"
date: "May 10, 2020"
output: html_document
---

```{r, echo = FALSE, warning = FALSE, message = FALSE}
# Packages! (oh boy there's going to be a lot of these)
library(tidyverse)
library(tidymodels)
library(ggplot2)
library(ggridges) # !!! remove if unused by end
```


```{r, echo = FALSE, warning = FALSE, message = FALSE}
# Data! (TidyTuesday 2020-03-10)
tuition_cost <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/tuition_cost.csv')

tuition_income <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/tuition_income.csv') 

salary_potential <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/salary_potential.csv')

historical_tuition <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/historical_tuition.csv')

diversity_school <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-03-10/diversity_school.csv')
```


MATH 228: Introduction to Data Science is a college course, and the work for this course (at the very least) must match the quality of work generally expected from an undergraduate-level student. For this project, you could say that a "college-level analysis" is required. What better data is there to perform a *"college-level analysis"* with than data *actually about colleges*!

You can find the TidyTuesday page for these data [here](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-03-10/readme.md).

## General research questions:

#  - Does one state have a higher avaerage tuition than any other states?
  - Is there an association between type of school (public, private, for-profit) and price of room and board?
  - Can we predict price of ---- from the available data?
#  - Are there any trends in the difference between total cost for tuition and net cost for tuition (average actually paid after scholarships/awards)?
#  - Is there an association between percent of students enrolled that are women and tuition?
 
  - Is tuition for colleges in Boston significantly different than tuition for colleges in Massachusetts?
  - seven sisters?

***

Let's start with this question: Does one state have a higher in-state tuition than any other states?

```{r}
tuition_cost %>% 
  filter(degree_length == "4 Year") %>% # only looking at 4 year colleges
  group_by(state) %>% 
  summarise(av_in_state_tu = mean(in_state_tuition)) %>% # creating variable for average in-state tuition for whole state
  na.omit() %>% # get rid of empty/NA state values
  ggplot(aes(x = fct_reorder(state, av_in_state_tu), y = av_in_state_tu)) + # fct_reorder() organizes bars in descending order (it looks WAY better!)
  geom_col() +
  theme_bw() + # my favorite theme right after theme_void()
  coord_flip() +
  labs(x = "", y = "", # x and y axis labels are redundant here
       title = "Average In-State Tuition by State",
       subtitle = "For 4-year degree programs",
       caption = "Source: Tuitiontracker.org") +
  scale_y_continuous(labels = scales::dollar_format()) # makes y axis labels into $ with commas
```

New England states are dominating the top spots for average in-state tuition. Rhode island has the most expensive average in-state tuition at \$36,114.90, and Massachusetts is a very close second at \$36,081.24 . Wyoming has a surprisingly low average in-state tuition at an even $5,400.

If this were a race, Rhode Island and Massachusetts would have a solid lead over the other 48 states - but this isn't a race, and I don't think anyone would actually want their state to win. 

*** 
Moving away from state data...


This might be a long shot, but all is fair in Data Exploration! Is there an association between percent of students enrolled that are women and tuition? Do women tend to go to pricey schools?

```{r}
tuition_cost %>% 
  left_join(diversity_school, by = c("name", "state")) %>% # join together tuition_income and diversity_school data
  filter(category == "Women") %>% 
  mutate(per_women = enrollment / total_enrollment) %>% # create a variable for percent of women enrolled
  ggplot(aes(x = per_women, y = in_state_tuition)) +
  geom_point() +
  geom_smooth(se = F) +
  facet_wrap(~ type, nrow = 3) +
  theme_bw() +
  labs(x = "Proportion Women Enrolled", y = "In-State Tuition",
       caption = "Data: Tuitiontracker.org & Chronicle of Higher Education", # tuition_cost from Tuitiontracker.org; diversity_school from Chron. Higher Ed
       title = "Do Women Prefer Pricy Colleges?") +
  scale_y_continuous(labels = scales::dollar_format()) + # make Y axis labels $
  scale_x_continuous(labels = scales::percent_format()) # make X axis labels %
 
```

There doesn't seem to be any clear association between percentage of women enrolled and tuition price. I did remember an interesting tid-bit while looking at this data visualization: "All boys" or "All women" schools are usually private schools, because single-sex public schools won't get financial aid from the government!
Specific information about the laws regarding this can be can be found [on the wikipedia page for single-sex education](https://en.wikipedia.org/wiki/Single-sex_education#United_States).

***

Are there any trends in the difference between total cost for tuition and net cost for tuition (average actually paid after scholarships/awards)?

In order to do this, I first filtered the data so we were only looking at data from 2018 for each school. I didn't need to keep net cost separated by income level, so I made an aggregate summary statistic for average net cost across all income levels (`av_net`) using `summarize()`. Since I can't use `mutate()` to create a new variable from an aggregate summary stat and a single data point, I made a second aggregate summary statistic for the average total price of tuition called `av_total` in the `summarize()` function (the values for `total_price` that I'm aggregating here should all be the same exact number so taking `mean(total_price)` shouldn't change anything). Now that they are both aggregate summary stats grouped by college name, I can use them both in a `mutate()` to create a variable representing the **average** percent of tuition actually paid. From there, I created a histogram to provide a visualization of the distribution of the average percent of tuition that is actually paid for each school.

```{r}
tuition_income %>% 
  filter(year == "2018") %>% #only looking at 2018 data to make this a little less complicated
  group_by(name) %>% 
  summarize(av_net = mean(net_cost), # create a variable for average net cost (for all income levels)
            av_total = mean(total_price)) %>% # need to do this to get aggregate summary stat of total cost
  mutate(per_paid = av_net / av_total) %>% 
  ggplot(aes(x = per_paid)) +
  geom_histogram(color = "white", fill = "chartreuse3") + # I had to put this color in somewhere :)
  theme_bw() +
  labs(x = "Percent of Tuition Actually Paid", y = "Number of Schools",
       caption = "Data: Tuitiontracker.org",
       title = "Average Percent of Tuition Paid Out of Pocket") +
  scale_x_continuous(labels = scales::percent_format()) # make X axis labels %
```

According to this histogram, the most common percentages of tuition actually paid is centered around 50%. The mean percent of tuition actually paid is 49.04%, with a standard deviation of 19.23%. The median percent of tuition actually paid is 50.71%. The number of schools with an average of 0% tuition actually paid is very surprising, and is probably due to missing values or data entry errors, as it does not make sense. This is troubling to see, as it raises questions about the validity of the data. 


The above data visualization uses the *average* net cost *for all income levels*. What if we keep the data separated by income level?

```{r}
tuition_income %>% 
  filter(year == "2018") %>% # still just focusing on 2018 data
  mutate(per_paid = net_cost / total_price) %>% 
  ggplot(aes(x = per_paid, y = income_lvl)) +
  geom_density_ridges() +
  labs(x = "Percent of Tuition Actually Paid", y = "Income Level",
       caption = "Data: Tuitiontracker.org",
       title = "Percent of Tuition Paid by Income Level") +
  scale_x_continuous(labels = scales::percent_format())
```

Percent of tuition actually paid seems to increase as income level increases, suggesting that financial aid is possibly given based on need. 


What if we also separated this based on school type (Public, Private, For-Profit)?

In order to do this, we have to join together the `tuition_income` data with the `tuition_cost` data:

```{r}
tuition_cost_short = tuition_cost %>% 
  select(name, state, state_code, type)

tuition_new = tuition_income %>% 
  rename(state_code = state) %>% 
  inner_join(tuition_cost_short, by = c("name", "state_code")) %>% 
  mutate(per_paid = net_cost / total_price)
  
ggplot(tuition_new, aes(x = per_paid, y = income_lvl)) +
  geom_density_ridges() +
  facet_wrap(~ type, nrow = 3) +
  labs(x = "Percent of Tuition Actually Paid", y = "Income Level",
       caption = "Data: Tuitiontracker.org",
       title = "Percent of Tuition Paid by Income Level and School Type") +
  scale_x_continuous(labels = scales::percent_format())
```

```{r, eval=FALSE, echo=FALSE}
# Optional 3 different graphs, if the facet_wrap is too small
tuition_new %>% 
  filter(type == "For Profit") %>% 
  ggplot(aes(x = per_paid, y = income_lvl)) +
  geom_density_ridges() +
  labs(x = "Percent of Tuition Actually Paid", y = "Income Level",
       caption = "Data: Tuitiontracker.org",
       title = "Percent of Tuition Paid by Income Level",
       subtitle = "For Profit Colleges Only") +
  scale_x_continuous(labels = scales::percent_format())

tuition_new %>% 
  filter(type == "Private") %>% 
  ggplot(aes(x = per_paid, y = income_lvl)) +
  geom_density_ridges() +
  labs(x = "Percent of Tuition Actually Paid", y = "Income Level",
       caption = "Data: Tuitiontracker.org",
       title = "Percent of Tuition Paid by Income Level",
       subtitle = "Private Colleges Only") +
  scale_x_continuous(labels = scales::percent_format())

tuition_new %>% 
  filter(type == "Public") %>% 
  ggplot(aes(x = per_paid, y = income_lvl)) +
  geom_density_ridges() +
  labs(x = "Percent of Tuition Actually Paid", y = "Income Level",
       caption = "Data: Tuitiontracker.org",
       title = "Percent of Tuition Paid by Income Level",
       subtitle = "Public Colleges Only") +
  scale_x_continuous(labels = scales::percent_format())
```





